---
title: "How Does the World Listen to Music?"
subtitle: "An exploration of how Global and USA music trends compare"
author: "Jude Groff"
institute: "Houghton University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader: 
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringan)
library(xaringanthemer)
library(kableExtra)
library(xaringanExtra)
library(dplyr)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")

knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```


```{r load-data, include=FALSE}
GMSLP <- read.csv(file = "../data/Global_Music_Streaming_Listener_Preferences.csv")
```
```{r}
GMSLP <- GMSLP %>%
  mutate(Region = ifelse(Country == "USA", "USA", "Non-USA"))

platform_counts <- GMSLP %>%
  group_by(Region, Streaming.Platform) %>%
  summarise(User_Count = n(), .groups = "drop")

region_totals <- platform_counts %>%
  group_by(Region) %>%
  summarise(Total_Users = sum(User_Count), .groups = "drop")

platform_usage <- platform_counts %>%
  left_join(region_totals, by = "Region") %>%
  mutate(Percentage = round((User_Count / Total_Users) * 100, 1))

top_platforms <- GMSLP %>%
  count(Streaming.Platform, sort = TRUE) %>%
  head(7) %>%
  pull(Streaming.Platform)

platform_usage <- platform_usage %>%
  mutate(Platform = ifelse(Streaming.Platform %in% top_platforms, 
                          as.character(Streaming.Platform), 
                          "Other"))

platform_usage <- platform_usage %>%
  group_by(Region, Platform) %>%
  summarise(User_Count = sum(User_Count),
            Total_Users = first(Total_Users),
            Percentage = round((User_Count / Total_Users) * 100, 1),
            .groups = "drop")

platform_order <- platform_usage %>%
  filter(Region == "Non-USA") %>%
  arrange(desc(Percentage)) %>%
  pull(Platform)

platform_usage$Platform <- factor(platform_usage$Platform, levels = platform_order)
```
```{r}
platform_percentages <- GMSLP %>%
  count(Streaming.Platform) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  arrange(desc(Percentage))

```

```{r, include=FALSE}
style_xaringan(
  title_slide_background_image = "https://images.unsplash.com/photo-1470225620780-dba8ba36b745?q=80&w=2070&auto=format&fit=crop&ixlib=rb-4.0.3&ixid=M3wxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8fA%3D%3D"
)
```


class: center, middle

##What do music trends look like globally? How does it compare to the US?
---

class: center, middle

##Streaming platforms are a huge part of global trends, as these have become the main source of music in the world. Globally, what music platform is the most used? 

---

class: center, middle

```{r}
GMSLP %>%
  group_by(Streaming.Platform) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = Count / sum(Count) * 100) %>%
  ungroup() %>%
  arrange(desc(Percentage))%>%
  select(-Count) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```


**Amazon music is the most used with the highest percentage of 17.22%**
---

class: center, middle

##But how does this compare to the US?

```{r}
GMSLP %>%
  filter(Country == "USA") %>%
  group_by(Streaming.Platform) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = Count / sum(Count) * 100) %>%
  ungroup() %>%
  arrange(desc(Percentage))%>%
  select(-Count) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```
---

class: center, middle

```{r}
ggplot(platform_usage, aes(x = Platform, y = Percentage, fill = Region)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), width = 0.8) +
  geom_text(aes(label = paste0(Percentage, "%")), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, 
            size = 3) +
  scale_fill_manual(values = c("USA" = "#E41A1C", "Non-USA" = "#377EB8"),
                    labels = c("Global Users (excluding USA)", "USA")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(title = "Streaming Platforms: USA vs. Global(excluding USA)",
       x = "Music Platform",
       y = "Percentage of Users (%)",
       fill = "User Location") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(color = "gray40"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )
```
---
class: center, middle

##What does Global vs. USA streaming trends look like when it comes to purchasing music subscriptions?
---
```{r}
GMSLP %>%
  count(Region, Subscription.Type) %>%
  group_by(Region) %>%
  mutate(Percentage = n / sum(n) * 100) %>%
  ungroup()%>%
  ggplot(aes(x = Region, y = n, fill = Subscription.Type)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = paste0(round(Percentage), "%")), 
            position = position_stack(vjust = 0.5),
            color = "white", fontface = "bold") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Subscription Types by Region",
    subtitle = "Distribution and percentages across geographic regions",
    x = "Region",
    y = "Count",
    fill = "Subscription Type"
  ) +
  theme_minimal()
```
---
class: center, middle

##Does this result in more listening time?

---
class: center, middle

**Average listening time Global**

```{r}
GMSLP %>%
  summarize(Average_Listening_Time_Global = mean(Minutes.Streamed.Per.Day, na.rm = TRUE)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F) 
```

**Average listening time USA**
```{r}
GMSLP %>%
  filter(Region == "USA") %>%
  group_by(Country) %>%
  summarize(Average_Listening_Time_USA = mean(Minutes.Streamed.Per.Day, na.rm = TRUE))%>%
  select(-Country)%>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```
---

class: center, middle

#Conclusion